---
title: "White House Wallet"
format: 
  dashboard:
    theme: flatly
    nav-buttons:
      - icon: github
        href: https://github.com/louiseoh18/project04-goldmagikarp.git
server: shiny
---

```{r}
#| context: setup
library(here)
library(shiny)
library(lubridate)
library(ggplot2)
library(tidyverse)

# Name ideas: PocketbookPresident, WalletWatch, ApprovalIndex
```

# Home

**Welcome to White House Wallet!**\n

This dashboard is designed to provide an accessible overview of historical presidential approval ratings, and explore how different economic indicators may play a role in a leader's public perception. The dashboard is interactive and user-friendly, allowing users to select different presidents, toggle between economic indicators of interest, and even modify values of economic indicators to predict how they will influence the approval rating of the current U.S. president. 

**Tutorial**\n

1. **Historical Data**
    - Select a U.S. president you would like to examine
    - The graph immediately below your selection will display all available approval ratings over the course of the president's service.
    - On the bottom graph, you can toggle between economic indicators which will display how those indicators changed over the course of the president's service.
    - Since both the approval rating graph and indicator graph span the same time frame, you can directly compare the two.
    - See how approval ratings and economic indicators were different between various presidents.

2. **Prediction and Modeling**
    - The first graph displays our pre-trained model's prediction for the current president's approval rating for the next 3-6 months. This model was trained based on the historical approval ratings and economic indicators from the dashboard's previous page.
    - The second graph...

# Historical Data

```{r}
#| context: setup

load(here("data/complete_data.rda"))

```

```{r}
#| context: server
econ_labels <- c(
  "Unemployment Rate" = "unemployment_rate",
  "Consumer Price Index" = "consumer_p_index",
  "Total Jobs" = "total_nonfarm_jobs",
  "Average Hourly Earnings" = "avg_hourly_earnings",
  "Real GDP" = "Real_GDP",
  "Disposable Income" = "Disposable_Income",
  "Corporate Profits" = "Corporate_Profits",
  "Healthcare Price Index" = "Healthcare_Price_Index",
  "Government Spending" = "Gov_Spending",
  "Personal Savings Rate" = "Personal_Saving_Rate"
)
pres_data <- reactive({
    final_combined_data |> filter(president == input$president)
  })
  
  # Approval rating plot
output$approval_plot <- renderPlot({
  pres_data() |>
    ggplot(aes(x = date)) +
    geom_line(aes(y = approval_rating, color = "Approval")) +
    geom_line(aes(y = disapproval_rating, color = "Disapproval")) +
    geom_line(aes(y = unsure_rating, color = "Unsure")) +
    scale_color_manual(values = c("#61D04F", "#DF536B", "#2297E6")) +
    labs(y = "Rating (%)", x = "Date", color = "Legend") +
    theme_minimal()
  })

output$econ_plot <- renderPlot({
    var <- input$econ_var
    friendly_label <- names(econ_labels)[econ_labels == var]
    pres_data() |>
      ggplot(aes(x = date, y = .data[[var]])) +
      geom_line(color = "mediumpurple1") +
      labs(y = friendly_label, x = "Date") +
      theme_minimal()
  })
```

## {.toolbar}

```{r}
selectInput(
        inputId = "president",
        label = "Choose a President:",
        choices = rev(unique(final_combined_data$president))
        [!is.na(unique(final_combined_data$president))],
        selected = "Trump (2nd Term)",
        selectize = FALSE
      )
```

## Approval Rating Card

```{r}
#| title: "Presidential Approval Ratings"
plotOutput("approval_plot")
```

## Economic Indicator Card

### Econ selection {width="20%"}

```{r}
econ_labels <- c(
  "Real GDP" = "Real_GDP",
  "Disposable Income" = "Disposable_Income",
  "Corporate Profits" = "Corporate_Profits",
  "Healthcare Price Index" = "Healthcare_Price_Index",
  "Government Spending" = "Gov_Spending",
  "Personal Savings Rate" = "Personal_Saving_Rate",
  "Unemployment Rate" = "unemployment_rate",
  "Consumer Price Index" = "consumer_p_index",
  "Total Jobs" = "total_nonfarm_jobs",
  "Average Hourly Earnings" = "avg_hourly_earnings"
)
selectInput(inputId = "econ_var",
            label = "Select Economic Indicator:",
            choices = econ_labels
            )
```

### Economic Plot {width = "80%"}

```{r}
#| title: "Economic Indicators"
plotOutput("econ_plot")

```

# Prediction and Modeling

## Trump's 6-Month Forecast

::: {.card title = "Trumps 6-Month Approval Forecast"}
```{r}

```
:::

## Simulation

:::{.card title = "Interactive Simulation"}
```{r}

```
:::

# References

