---
title: "U.S. Presidential Approval Ratings"
format: 
  dashboard:
    theme: flatly
    nav-buttons:
      - icon: github
        href: https://github.com/louiseoh18/project04-goldmagikarp.git
server: shiny
---

```{r}
#| context: setup
library(shiny)
library(ggplot2)
library(tidyverse)
```

# Home

**Welcome to [name of dashboard]!**\n

This dashboard is designed to provide an accessible overview of historical presidential approval ratings, and explore how different economic indicators may play a role in a leader's public perception. The dashboard is interactive and user-friendly, allowing users to select different presidents, toggle between economic indicators of interest, and even modify values of economic indicators to predict how they will influence the approval rating of the current U.S. president. 

**Tutorial**\n

1. **Historical Data**
    - Select a U.S. president you would like to examine
    - The graph immediately below your selection will display all available approval ratings over the course of the president's service.
    - On the bottom graph, you can toggle between economic indicators which will display how those indicators changed over the course of the president's service.
    - Since both the approval rating graph and indicator graph span the same time frame, you can directly compare the two.
    - See how approval ratings and economic indicators were different between various presidents.

2. **Prediction and Modeling**
    - The first graph displays our pre-trained model's prediction for the current president's approval rating for the next 3-6 months. This model was trained based on the historical approval ratings and economic indicators from the dashboard's previous page.
    - The second graph...

# Historical Data

## Selection Card

::: {.card title = "Choose a President:"}
```{r}
# selectInput(inputID = "president", 
#             label ="President", 
#             choices = names("president"),
#             selected = "")

```
:::

## Approval Rating Card

::: {.card title = "Presidential Approval Ratings"}
```{r}

```
:::

## Economic Indicator Card {.tabset}

::: {.card title = "Unemployment Rate"}
```{r}

```
:::

::: {.card title = "Consumer Price Index"}
```{r}

```
:::

::: {.card title = "Total Jobs"}
```{r}

```
:::

::: {.card title = "Average Hourly Wages"}
```{r}

```
:::

::: {.card title = "Real GDP"}
```{r}

```
:::

::: {.card title = "Disposable Income"}
```{r}

```
:::

::: {.card title = "Corporate Profits"}
```{r}

```
:::

::: {.card title = "Healthcare Price Index"}
```{r}

```
:::

::: {.card title = "Government Spending"}
```{r}

```
:::

::: {.card title = "Personal Savings Rate"}
```{r}

```
:::

# Prediction and Modeling

## Trump's 6-Month Forecast

::: {.card title = "Trumps 6-Month Approval Forecast"}
```{r}

```
:::

## Simulation

:::{.card title = "Interactive Simulation"}
```{r}

```
:::

# References

