---
title: "White House Wallet"
format: 
  dashboard:
    theme: flatly
    nav-buttons:
      - icon: github
        href: https://github.com/louiseoh18/project04-goldmagikarp.git
server: shiny
bibliography: my-refs.bib
nocite: |
  @*
---

```{r}
#| context: setup
library(here)
library(shiny)
library(lubridate)
library(ggplot2)
library(tidyverse)

# Name ideas: PocketbookPresident, WalletWatch, ApprovalIndex, 
# ExecutiveEconomics, Parralax View

# Another note: we need graphics for all pages (namely home page, historical data, and references)

# Reference page so maybe describe how the different data was used. We didn't end up using some of the data in the actual dashboard but we still analyzed it (which viewers won't see but can be made aware of)
```

# Home

## Welcome message

**Welcome to White House Wallet!**\n

This dashboard is designed to provide an accessible overview of historical presidential approval ratings, and explore how different economic indicators may play a role in a leader's public perception. The dashboard is interactive and user-friendly, allowing users to select different presidents, toggle between economic indicators of interest, and even modify values of economic indicators to predict how they will influence the approval rating of the current U.S. president. 

**TUTORIAL**\n

1. **Historical Data**
    - Select a U.S. president you would like to examine
    - The graph immediately below your selection will display all available approval ratings over the course of the president's service.
    - On the bottom graph, you can toggle between economic indicators which will display how those indicators changed over the course of the president's service.
    - Since both the approval rating graph and indicator graph span the same time frame, you can directly compare the two.
    - See how approval ratings and economic indicators were different between various presidents.
    
2. **Comparing Presidents**
    - Select two U.S. presidents to compare. Use Ctrl (Windows) or Command (Mac) to select the second president.
    - The graph below will display approval ratings for both presidents over their time in office (measured in months).
    - The solid line represents the first selected president, and the dotted line represents the second.
    - Because both lines appear on the same graph, you can easily compare differences in approval ratings between the two leaders.

3. **Prediction and Modeling**
    - The graph displays our pre-trained model's prediction for the current president's approval rating for the next 6 months (status quo). This model was trained based on the historical approval ratings and economic indicators from the dashboard's previous page.
    - The dotted blue line shows the status quo which is predicted approval rating assuming most recent economic data stays consistent indefinitely.
    - The red line shows the simulated prediction based on your own "what-if" scenarios. You can modify the economic indicators in the side panel to the left to see how each one may change the president's approval rating.
    - You can see summaries of these predicted approval ratings in the right-most panels for the status quo and the simulation (6-month and long-term predictions).

# Historical Data

```{r}
#| context: setup
load(here("data/complete_data.rda"))
```

```{r}
#| context: server
econ_labels <- c(
  "Real GDP" = "Real_GDP",
  "Disposable Income" = "Disposable_Income",
  "Corporate Profits" = "Corporate_Profits",
  "Healthcare Price Index" = "Healthcare_Price_Index",
  "Government Spending" = "Gov_Spending",
  "Personal Savings Rate" = "Personal_Saving_Rate",
  "Unemployment Rate" = "unemployment_rate",
  "Consumer Price Index" = "consumer_p_index",
  "Total Jobs" = "total_nonfarm_jobs",
  "Average Hourly Earnings" = "avg_hourly_earnings"
)
pres_data <- reactive({
    final_combined_data |> filter(president == input$president)
  })
  
# Approval rating plot
output$approval_plot <- renderPlot({
  pres_data() |>
    ggplot(aes(x = date)) +
    geom_line(aes(y = approval_rating, color = "Approval"), size = 1) +
    geom_line(aes(y = disapproval_rating, color = "Disapproval"), size = 1) +
    geom_line(aes(y = unsure_rating, color = "Unsure"), size = 1) +
    scale_color_manual(values = c("#61D04F", "#DF536B", "#2297E6")) +
    labs(y = "Rating (%)", x = "Date", color = "Legend") +
    theme_minimal()
  })

# Economic indicator plot
output$econ_plot <- renderPlot({
    var <- input$econ_var
    friendly_label <- names(econ_labels)[econ_labels == var]
    pres_data() |>
      ggplot(aes(x = date, y = .data[[var]])) +
      geom_line(color = "mediumpurple", size = 1) +
      labs(y = friendly_label, x = "Date") +
      theme_minimal()
  })
```

## {.toolbar}

```{r}
selectInput(
        inputId = "president",
        label = "Choose a President:",
        choices = rev(unique(final_combined_data$president)
        [!is.na(unique(final_combined_data$president))]),
        selected = "Trump (2nd Term)",
        selectize = FALSE
      )
```

## Approval Rating Card

### Historical page graphic {width="20%"}

```{r}
h4("filler for graphic")
```

### Approval plot {width="80%"}

```{r}
#| title: "U.S. Presidential Approval Ratings"
plotOutput("approval_plot")
```

## Economic Indicator Card

### Econ selection {width="20%"}

```{r}
econ_labels <- c(
  "Real GDP" = "Real_GDP",
  "Disposable Income" = "Disposable_Income",
  "Corporate Profits" = "Corporate_Profits",
  "Healthcare Price Index" = "Healthcare_Price_Index",
  "Government Spending" = "Gov_Spending",
  "Personal Savings Rate" = "Personal_Saving_Rate",
  "Unemployment Rate" = "unemployment_rate",
  "Consumer Price Index" = "consumer_p_index",
  "Total Jobs" = "total_nonfarm_jobs",
  "Average Hourly Earnings" = "avg_hourly_earnings"
)
selectInput(inputId = "econ_var",
            label = "Select Economic Indicator:",
            choices = econ_labels
            )
```

### Economic Plot {width = "80%"}

```{r}
#| title: "Economic Indicators"
plotOutput("econ_plot")

```

# Comparing Presidents

```{r}
#| context: setup

approval_comparison_data <-
  final_combined_data |>
  drop_na(1:5) |>
  group_by(president) |>
  mutate(
    months_in_office = case_when(
      month(min(date)) == 1 ~ interval(min(date), date) %/% months(1),
      month(min(date)) == 2 ~ interval(min(date), date) %/%
        months(1) +
        (month(min(date)) - 1), # presi's missing first month
      month(min(date)) == 8 ~ interval(min(date), date) %/%
        months(1) +
        (month(min(date)) - 8), # Ford
      month(min(date)) == 6 ~ interval(min(date), date) %/%
        months(1) +
        (month(min(date)) - 4), # Truman
      month(min(date)) == 12 ~ interval(min(date), date) %/%
        months(1) +
        (month(min(date)) - 11), # LBJ
      month(min(date)) == 7 ~ interval(min(date), date) %/%
        months(1) +
        (month(min(date)) - 1), # FDR 3rd term
    ),
    president = case_when(
      str_detect(president, "Trump") ~ "Donald Trump",
      TRUE ~ president
    ),
    months_in_office = case_when(
      date < "2025-01-01" ~ months_in_office,
      president == "Donald Trump" & date >= "2025-01-01" ~ months_in_office +
        49,
      president == "Joe Biden" & date == "2025-01-01" ~ 48
    )
  ) |>
  select(1:5, 17)
#will move to data folder later, currently just seeing if it works in dashboard
```

```{r}
#| context: server

presi_data <- 
  reactive({
    approval_comparison_data |> filter(president %in% input$president)
  })

output$comparison_approval_plot <- renderPlot({
  presi_data() |>
    ggplot(aes(x = months_in_office, linetype = president)) +
    geom_line(aes(y = approval_rating, color = "Approval"), size = 1) +
    geom_line(aes(y = disapproval_rating, color = "Disapproval"), size = 1) +
    geom_line(aes(y = unsure_rating, color = "Unsure"), size = 1) +
    scale_color_manual(values = c("#61D04F", "#DF536B", "#2297E6")) +
    scale_linetype_manual(values = 
                            c("solid", "dashed", "dotted", "dotdash", "longdash", "twodash",
                              "12", "21", "13", "31", "14", "41", "1122", "2211", "F2")) +
      labs(y = "Rating (%)", x = "Time in Office (months)", color = "Legend") +
      theme_minimal()
  })
```

```{r}
selectInput(
  inputId = "president",
  label = "Choose Presidents",
  choices = replace(
    rev(unique(approval_comparison_data$president)),
    c(1, 2),
    rev(unique(approval_comparison_data$president))[c(2, 1)]
    ),
  selected = c("Donald Trump", "Joe Biden"),
  multiple = TRUE,
  selectize = FALSE
  )
```

```{r}
# plot outputs
plotOutput("comparison_approval_plot")
```

# Prediction and Modeling

```{r}
#| context: setup

library(shiny)
library(dplyr)
library(ggplot2)
library(lubridate)
library(randomForest)
library(here)
library(zoo)


rf_model <- readRDS(here("data/rf_approval_model.rds"))
trump_history <- readRDS(here("data/trump_history.rds"))


last_actual_row <- tail(trump_history, 1)
start_date <- last_actual_row$date
start_approval <- last_actual_row$approval_rating

last_3_approvals <- tail(trump_history$approval_rating, 3)
if(length(last_3_approvals) < 3) {
  last_3_approvals <- rep(start_approval, 3)
}

```

```{r}
#| context: server

observeEvent(input$reset, {
    updateNumericInput(session, "unemp_start", value = 4.5)
    updateNumericInput(session, "unemp_end", value = 5.0)
    updateNumericInput(session, "inf_start", value = 3.0)
    updateNumericInput(session, "inf_end", value = 2.5)
    updateNumericInput(session, "gdp_start", value = 2.5)
    updateNumericInput(session, "gdp_end", value = 1.8)
    updateSliderInput(session, "wages", value = 3.5)
    updateSliderInput(session, "healthcare", value = 6.0)
    updateSliderInput(session, "jobs", value = 1.0)
    updateSliderInput(session, "market", value = 4.0)
  })
  
  forecast_data <- reactive({
    future_dates <- seq(start_date + months(1), as.Date("2027-12-01"), by = "month")
    n_months <- length(future_dates)
    
    unemp_seq <- seq(input$unemp_start, input$unemp_end, length.out = n_months)
    inf_seq   <- seq(input$inf_start, input$inf_end, length.out = n_months)
    gdp_seq   <- seq(input$gdp_start, input$gdp_end, length.out = n_months)
    
    df <- data.frame(
      date = future_dates,
      president = "Trump (2nd Term)", 
      Party = "Republican",
      Year_Index = year(future_dates),
      Months_in_Office = seq(11, 11 + n_months - 1),
      Honeymoon = 0,
      
      Unemployment      = unemp_seq,
      Inflation_Rate    = inf_seq,
      GDP_Growth        = gdp_seq,
      Wage_Growth       = rep(input$wages, n_months),
      Healthcare_Infl   = rep(input$healthcare, n_months),
      Job_Growth        = rep(input$jobs, n_months),
      Profit_Growth     = rep(input$market, n_months),
      Income_Growth     = rep(2.0, n_months), 
      Gov_Spend_Growth  = rep(1.0, n_months),
      Savings_Rate      = rep(4.0, n_months),
      
      Approval_Lag3     = NA,
      approval_rating   = NA
    )
    
    approval_buffer <- last_3_approvals
    
    for(i in 1:n_months) {
      df$Approval_Lag3[i] <- approval_buffer[1]
      pred_val <- predict(rf_model, newdata = df[i, ])
      df$approval_rating[i] <- pred_val
      approval_buffer <- c(approval_buffer[-1], pred_val)
    }
    
    first_pred <- df$approval_rating[1]
    anchor_gap <- start_approval - first_pred
    df$approval_rating <- df$approval_rating + anchor_gap
    
    df$Type <- "Custom Scenario"
    return(df)
  })
  
  baseline_forecast <- reactive({
    future_dates <- seq(start_date + months(1), as.Date("2027-12-01"), by = "month")
    n_months <- length(future_dates)
    
    df <- data.frame(
      date = future_dates,
      president = "Trump (2nd Term)", 
      Party = "Republican",
      Year_Index = year(future_dates),
      Months_in_Office = seq(11, 11 + n_months - 1),
      Honeymoon = 0,
      
      Unemployment      = rep(4.5, n_months),
      Inflation_Rate    = rep(3.1, n_months),
      GDP_Growth        = rep(2.3, n_months),
      Wage_Growth       = rep(4.0, n_months),
      Healthcare_Infl   = rep(5.5, n_months),
      Job_Growth        = rep(1.1, n_months),
      Profit_Growth     = rep(4.5, n_months),
      Income_Growth     = rep(2.0, n_months), 
      Gov_Spend_Growth  = rep(1.0, n_months),
      Savings_Rate      = rep(4.5, n_months),
      
      Approval_Lag3     = NA,
      approval_rating   = NA
    )
    
    approval_buffer <- last_3_approvals
    
    for(i in 1:n_months) {
      df$Approval_Lag3[i] <- approval_buffer[1]
      pred_val <- predict(rf_model, newdata = df[i, ])
      df$approval_rating[i] <- pred_val
      approval_buffer <- c(approval_buffer[-1], pred_val)
    }
    
    first_pred <- df$approval_rating[1]
    anchor_gap <- start_approval - first_pred
    df$approval_rating <- df$approval_rating + anchor_gap
    
    df$Type <- "Status Quo"
    return(df)
  })
  
  output$forecastPlot <- renderPlot({
    fc_custom <- forecast_data()
    fc_base <- baseline_forecast()
    
    hist_plot <- trump_history %>% mutate(Type = "Actual History")
    bridge_custom <- last_actual_row; bridge_custom$Type <- "Custom Scenario"
    bridge_base <- last_actual_row; bridge_base$Type <- "Status Quo"
    
    plot_df <- bind_rows(
      hist_plot, 
      bridge_custom, fc_custom %>% select(date, approval_rating, Type),
      bridge_base, fc_base %>% select(date, approval_rating, Type)
    )
    
    six_month_date <- start_date + months(6)
    
    ggplot(plot_df, aes(x = date, y = approval_rating, color = Type, linetype = Type, size = Type)) +
      geom_line() +
      geom_point(data = filter(plot_df, Type == "Actual History"), size = 3, alpha = 0.8) +
      
      geom_vline(xintercept = start_date, linetype = "dotted", color = "gray50") +
      geom_vline(xintercept = six_month_date, linetype = "dashed", color = "#f39c12", alpha = 0.6) +
      annotate("text", x = six_month_date, y = max(plot_df$approval_rating, na.rm=T), 
               label = "6 Months", angle = 90, vjust = -1, color = "#f39c12") +
      
      scale_color_manual(values = c("Actual History" = "black", "Custom Scenario" = "#D55E00", "Status Quo" = "#2980b9")) +
      scale_linetype_manual(values = c("Actual History" = "solid", "Custom Scenario" = "dashed", "Status Quo" = "dotted")) +
      scale_size_manual(values = c("Actual History" = 1.5, "Custom Scenario" = 1.5, "Status Quo" = 1.2)) +
      
      labs(y = "Approval Rating (%)", x = "Year") +
      theme_minimal(base_size = 16) +
      theme(legend.position = "bottom") +
      coord_cartesian(ylim = c(25, 60))
  })
  
  output$six_month_score <- renderText({
    val <- forecast_data()$approval_rating[6]
    paste0(round(val, 1), "%")
  })
  
  output$status_quo_score <- renderText({
    val <- baseline_forecast()$approval_rating[6]
    paste0(round(val, 1), "%")
  })
  
  output$final_score <- renderText({
    val <- tail(forecast_data()$approval_rating, 1)
    paste0(round(val, 1), "%")
  })
  
  output$change_score <- renderText({
    start_val <- start_approval
    end_val <- tail(forecast_data()$approval_rating, 1)
    diff <- end_val - start_val
    sign <- ifelse(diff >= 0, "+", "")
    paste0("(", sign, round(diff, 1), "% from today)")
  })

```

## {.sidebar}

### Economic Trajectory

Define the start and end points. The model simulates the trend between them.

```{r}

hr()

# "1. Unemployment Rate (%)"
div(
  h6("Unemployment Rate (%)"),
  splitLayout(
    numericInput("unemp_start", "Start (Jan '26)", value = 4.5, step = 0.1),
    numericInput("unemp_end", "End (Dec '27)", value = 5.0, step = 0.1)
  )
)


# "2. Inflation Rate (Year-over-Year %)"
div(
  h6("Inflation Rate (Year-over-Year %)"),
      splitLayout(
        numericInput("inf_start", "Start", value = 3.0, step = 0.1),
        numericInput("inf_end", "End", value = 2.5, step = 0.1)
      )
)

# "3. GDP Growth (Year-over-Year %)"
div(
  h6("GDP Growth (Year-over-Year %)"),
      splitLayout(
        numericInput("gdp_start", "Start", value = 2.5, step = 0.1),
        numericInput("gdp_end", "End", value = 1.8, step = 0.1)
      ),
)

hr()

# "Secondary Factors (Avg)")
h6("Secondary Factors (Avg)")
sliderInput("wages", "Wage Growth (%):", min = 0, max = 6, value = 3.5, step = 0.1)
sliderInput("healthcare", "Healthcare Inflation (%):", min = 0, max = 10, value = 6.0, step = 0.5)
sliderInput("jobs", "Job Growth (%):", min = -2, max = 4, value = 1.0, step = 0.1)
sliderInput("market", "Corp. Profit Growth (%):", min = -10, max = 15, value = 4.0, step = 1)

actionButton("reset", "Reset Scenarios", class = "btn-secondary", width = "100%")
```

## Main panels {width="80%"}

### Plot {height="75%"}
```{r}
#| title: "Presidential Approval Forecaster: Trump 2nd Term"
plotOutput("forecastPlot", height = "500px")
```

### Model logic {height="25%"}
```{r}
h5("Model Logic:")
tags$ul(
  tags$li("Blue Line: 'Status Quo' - Assumes current economic conditions persist indefinitely."),
  tags$li("Red Line: 'Scenario' - Your custom forecast based on the sliders."),
  tags$li("Anchoring: Both forecasts start at the last known approval rating (36%).")
  )
```

## Other Cards {width="20%"}

### card 1 {height="40%"}
```{r}
div(class = "alert",
    style = "background-color: #2980b9;",
    h6("Status Quo (May '26)\n", 
       span(textOutput("status_quo_score", inline = TRUE), style = "font-weight: bold; font-size: 2.0em;")),
    p("Prediction if economy stays exactly as it is today.", style = "color: black; font-size: 0.9em;"))
```

### card 2 {height="60%"}
```{r}
div(class = "alert",
    style = "background-color: #D55E00;",
    h6("Scenario (May '26)\n", 
       span(textOutput("six_month_score", inline = TRUE), style = "font-weight: bold; font-size: 2.0em;")),
    p("Prediction based on your inputs.", style = "color:black; font-size: 0.9em;"),
    h6("\n\n\nLong-Term (Dec '27)\n", 
       span(textOutput("final_score", inline = TRUE), style = "font-weight: bold; font-size: 2.0em;")),
    span(textOutput("change_score"), style = "font-size: 0.9em; color: black;"))
```

# References

The data was pulled from a variety of databases using API's

::: {#refs}
:::